Downward shortwave solar irradiance enhancement by clouds.

### 2024-04-22 article/article.tex
Commit 2024-04-22 06:14

After evaluating the modeled clear radiation for the different
atmospheric conditions, in relation to the measured GHI data, we choose
as a representative of the clear sky radiation, the case where
\(\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma\) and
\(w_{h\text{cs}} = w_h - 1\sigma\) (\(\text{GHI}_\text{CSm}\)). These
values represent a typical atmosphere in Thessaloniki with low load of
aerosols and humidity, which are the main factor that attenuate the GHI,
excluding clouds. To create the limit of CE identification, we created a
two branched threshold, as a function of SZA. A constant factor for low
SZAs, and a higher ramped factor for higher SZAs
(Equation\nobreakspace\ref{eq:CE4}). This is the criterion of our CE
identification. \begin{equation}
\text{CE} : \text{E}_\text{i} > \text{E}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{E}_\text{CSlim,i} = 1.05 \cdot \text{E}_\text{CSm,i}, & \text{$\theta \leq 60^\circ$}\\
\text{E}_\text{CSlim,i} = \left ({ 1.18 + \frac{1.05 - 1.18}{60 - 78} \cdot (\theta- 78) } \right ) \cdot \text{E}_\text{CSm,i}, & \text{$ 78^\circ > \theta > 60^\circ$}\\
\text{Excluded measurements}, & \theta > 78^\circ
\end{cases}\label{eq:CE4}
\end{equation} where: \(\theta\) is the solar zenith angle, \(\text{E}\)
the measured irradiance, \(\text{E}_\text{CSm}\) the selected modeled
clear sky irradiance, and \(i\) each of the one-minute observation.

@@

Our dataset, after the data selection processing, consists of 6144534
records of GHI, of which \(1.799\,\%\) are CE and \(0.036\,\%\) are ECE
events. The highest GHI recorder was \(1416.6\,W/m^2\) on 24~May 2007.
The absolute stronger CE event had an OIR of \(341.79\,W/m^2\) on 15~May
2014. The relative stronger CE event was \(54\,\%\) above the clear sky
threshold, on 28~October 2016.

@@

We computed the daily trend of the mean OIR of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order
autoregressive model with lag of 1 day, using the `maximum likelihood'
fitting method \citep{Gardner1980, Jones1980} by implementing the
function `arima' from the library `stats' of the R programming language
\citep{RCT2023}. The trends are reported together with the \(2\sigma\)
error. We observe an increase of \(0.216\pm 0.083\,W/m^2/y\) on the mean
OIR.

@@

Although, the previous result is closer to the raw data, we preferred to
present the annual statistics, that give a more clear picture about the
long term CE trends. Hence, we calculated the annual values from the
one-minute measurements. The trend of OIR is \(0.23\pm 0.11\,W/m^2/y\)
(Figure\nobreakspace{}\ref{fig:P-energy-mean}), but this value alone is
not very useful due the intrinsic high variability of this metric.

@@

A better indicator of changes on the characteristic of CE would be the
number of CE occurrences and the total energy of the CE over irradiance.
The annual number of CE occurrences, shows a steady increase of
\(122.2\pm 22.8\,\text{cases}/y\)
(Figure\nobreakspace{}\ref{fig:P-energy-N}). We have to note that the
energy related to the CE events can not be directly linked with the
total energy balance on the atmosphere. The net sun radiation of the
region is not increased, but rather redistributed through the CE.
Although, the instantaneous values, can exceed the equivalent clear sky
irradiance to a considerable level.

@@

Subsequently, the mean annual increase of energy due to CE events is
\(341.6\pm 73.1\,kJ/y\) (Figure~\ref{fig:P-energy-sum}), witch follows
the trend of the number of occurrences.

@@

In order to further study the characteristics of the CE events, we
grouped the single minute CE events, to continuous CE groups (CEG).
Thus, a CEG consists of one or more successive CE cases, and can have a
duration of multiple of minutes. We have identified 29009 groups, where
the group with the longest duration lasted 140 minutes on 07~July 2013.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can
conclude that longer durations are getting increasingly rare, with
durations above 10 minutes very rare.

>>>>

After evaluating the modeled clear radiation for the different
atmospheric conditions, in relation to the measured GHI data, we choose
as a representative of the clear sky radiation, the case where
\(\tau_{\text{cs}} = \tau_{500\text{nm}} - 1\sigma\) and
\(w_{h\text{cs}} = w_h - 1\sigma\) (\(\text{GHI}_\text{CSm}\)). These
values represent a typical atmosphere in Thessaloniki with low load of
aerosols and humidity, which are the main factor that attenuate the GHI,
excluding clouds. To create the limit of CE identification, we created a
two branched threshold, as a function of SZA. A constant factor for low
SZAs, and a higher ramped factor for higher SZAs
(Equation\nobreakspace\ref{eq:CE4}). This is the criterion of our CE
identification. \begin{equation}
\text{CE} : \text{E}_\text{i} > \text{E}_\text{CSlim,i}, \text{where} \begin{cases}
 \text{E}_\text{CSlim,i} = 1.04 \cdot \text{E}_\text{CSm,i}, & \text{$\theta \leq 90^\circ$}\\
\text{E}_\text{CSlim,i} = \left ({ 1.04 + \frac{1.04 - 1.04}{90 - 78} \cdot (\theta- 78) } \right ) \cdot \text{E}_\text{CSm,i}, & \text{$ 78^\circ > \theta > 90^\circ$}\\
\text{Excluded measurements}, & \theta > 78^\circ
\end{cases}\label{eq:CE4}
\end{equation} where: \(\theta\) is the solar zenith angle, \(\text{E}\)
the measured irradiance, \(\text{E}_\text{CSm}\) the selected modeled
clear sky irradiance, and \(i\) each of the one-minute observation.

@@

Our dataset, after the data selection processing, consists of 6144534
records of GHI, of which \(1.592\,\%\) are CE and \(0.036\,\%\) are ECE
events. The highest GHI recorder was \(1416.6\,W/m^2\) on 24~May 2007.
The absolute stronger CE event had an OIR of \(335.28\,W/m^2\) on 24~May
2007. The relative stronger CE event was \(50.2\,\%\) above the clear
sky threshold, on 28~October 2016.

@@

We computed the daily trend of the mean OIR of CE
(Figure\nobreakspace{}\ref{fig:CEmeanDaily}), using a first-order
autoregressive model with lag of 1 day, using the `maximum likelihood'
fitting method \citep{Gardner1980, Jones1980} by implementing the
function `arima' from the library `stats' of the R programming language
\citep{RCT2023}. The trends are reported together with the \(2\sigma\)
error. We observe an increase of \(0.202\pm 0.083\,W/m^2/y\) on the mean
OIR.

@@

Although, the previous result is closer to the raw data, we preferred to
present the annual statistics, that give a more clear picture about the
long term CE trends. Hence, we calculated the annual values from the
one-minute measurements. The trend of OIR is \(0.22\pm 0.11\,W/m^2/y\)
(Figure\nobreakspace{}\ref{fig:P-energy-mean}), but this value alone is
not very useful due the intrinsic high variability of this metric.

@@

A better indicator of changes on the characteristic of CE would be the
number of CE occurrences and the total energy of the CE over irradiance.
The annual number of CE occurrences, shows a steady increase of
\(110.1\pm 20.3\,\text{cases}/y\)
(Figure\nobreakspace{}\ref{fig:P-energy-N}). We have to note that the
energy related to the CE events can not be directly linked with the
total energy balance on the atmosphere. The net sun radiation of the
region is not increased, but rather redistributed through the CE.
Although, the instantaneous values, can exceed the equivalent clear sky
irradiance to a considerable level.

@@

Subsequently, the mean annual increase of energy due to CE events is
\(293.3\pm 63.1\,kJ/y\) (Figure~\ref{fig:P-energy-sum}), witch follows
the trend of the number of occurrences.

@@

In order to further study the characteristics of the CE events, we
grouped the single minute CE events, to continuous CE groups (CEG).
Thus, a CEG consists of one or more successive CE cases, and can have a
duration of multiple of minutes. We have identified 26042 groups, where
the group with the longest duration lasted 140 minutes on 07~July 2013.
By examining the frequency distribution of the CEG durations
(Figure\nobreakspace{}\ref{fig:ceg-duration-distribution}), we can
conclude that longer durations are getting increasingly rare, with
durations above 10 minutes very rare.
